---
title: "Penguins"
author: "Nabeelsoomro779"
date: "2024-01-11"
output:
  html_document: default
---
# Exploring the Palmer Penguins dataset

## Objectives:
1. Is there a correlation between flipper length and body mass?
2. Which penguin species is the largest?
3. Does sex affect body mass?


#### Environment Setup
###### Loading the following pacakges
* tidyverse
* ggplot2
* palmerpenguins
* gridExtra

```{r eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(palmerpenguins)
library(gridExtra)
```


The Penguins Dataset:-

```{r echo = FALSE}
head(penguins)
```

# Objective 1:
## Finding if there is a correlation between mass & flipper length.

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
plot1 = ggplot(data = penguins)+
  geom_point(mapping = aes(x = flipper_length_mm,
                           y = body_mass_g,
                           color = species))+
  labs(title = "Penguins: Body mass Vs. Flipper length",
       caption = "Data collected by Dr Kristen Gormen")
```

```{r warning=FALSE, echo=FALSE, message=FALSE, results='hide'}
plot2 = ggplot(data = penguins)+
  geom_smooth(mapping = aes(x = flipper_length_mm,
                            y = body_mass_g))+
  labs(title = "Positive correlation line")
```

```{r warning = FALSE, message=FALSE, echo=FALSE}
grid.arrange(plot1, plot2, ncol = 2, widths = c(1.6, 1))
```

> Postive correlation between the two variables.
as one variable changes, so does the other!


# Objective 2
## Finding out which species of penguins is the heaviest.

```{r warning=FALSE, echo=FALSE}
avg_mass_by_species <- penguins %>% 
  group_by(species) %>% 
  summarize(avg_mass = mean(body_mass_g, na.rm = TRUE))
```

```{r warning=FALSE, echo=FALSE}
ggplot(data = avg_mass_by_species)+
  geom_col(mapping = aes(x = species,
                         y = avg_mass,
                         fill = species, width = 0.6))+
  labs(title= "Mean mass by species",
       subtitle = "Gentoo is the largest species in terms of average mass")
```


# Objective 3
## Does sex play a role on weight?

```{r results='hide', warning=FALSE, message=FALSE, echo=FALSE}
avg_mass_by_species_and_sex <- penguins %>% 
  group_by(species, sex) %>% 
  summarize(avg_mass = mean(body_mass_g, na.rm = TRUE))
```

```{r echo=FALSE}
clean <- na.omit(avg_mass_by_species_and_sex)
```

```{r warning=FALSE, message=FALSE, echo=FALSE}
ggplot(data = clean)+
  geom_col(mapping = aes(x = species,
                         y = avg_mass,
                         fill = species, width = 0.6))+
  facet_wrap(~sex)+
  labs(title = "Mass difference between males & females",
       subtitle = "Males tend to have more mass than females")
```

